Fast approximate Bayesian inference of HIV indicators using PCA adaptive Gauss-Hermite quadrature

### 2023-07-22 src/docs_statml-poster/statml-poster.Rmd
Complete StatML poster #57

* Malawi is a relatively small country but still has latent field $\text{dim}(x) = 491$ and hyperparameters $\text{dim}(\theta) = 24$

@@

ks_reduction <- ks %>%
  filter(type == "Latent") %>%
  summarise(
    ks_tmb = mean(TMB),
    ks_aghq = mean(aghq)
  ) %>%
  mutate(reduction = 100 * (ks_aghq - ks_tmb) / ks_tmb) %>%
  pull(reduction) %>%
  round()

@@

* Run three methods: 1. TMB (baseline, `r round(time_taken$TMB, 0)` secs), 2. PCA-AGHQ (new, `r round(time_taken$aghq, 0)` hour), 3. NUTS (gold-standard, `r round(time_taken$tmbstan, 1)` days)
  * For PCA-AGHQ $k = 3$ and $s = 8$ chosen using Scree plot to explain ~90% of variance
  * For NUTS 4 chains of 100,000 thinned by 40 were required for good diagnostics e.g. all $\hat R < 1.025$
* Naomi can be used to assess probabilities targets have been met e.g. 90% of those who know their HIV status are on ART ("second 90")
  * Though PCA-AGHQ is better (`r round(second90_rmse, 0)`% RMSE), both TMB and PCA-AGHQ are biased (Figure 5)

```{r fig5, echo=FALSE, fig.align='center', fig.cap='PCA-AGHQ somewhat improves the second 90 bias.'}
knitr::include_graphics("fig5.png")
```

@@

* **Can we do any better than modest improvements?**
* Laplace marginals with matrix algebra approximations [@wood2020simplified] to speed up calculations
* Further methods for allocation of effort to "important" dimensions of hyperparameter grid

@@

**Contact** `r fontawesome::fa("home", fill = "black")` `athowes.github.io`; `r fontawesome::fa("envelope", fill = "black")` `ath19@ic.ac.uk`

>>>>

* Malawi is a relatively small country but still has latent field $\text{dim}(x) = 467$ and hyperparameters $\text{dim}(\theta) = 24$

@@

ks_summary <- ks %>%
  filter(type == "Latent") %>%
  group_by(type) %>%
  summarise(
    ks_tmb = mean(size * TMB) / sum(size),
    ks_aghq = mean(size * aghq) / sum(size)
  ) %>%
  mutate(reduction = 100 * (ks_aghq - ks_tmb) / ks_tmb) %>%
  pull(reduction) %>%
  round()

@@

* Run three methods: 1. TMB (baseline, `r round(time_taken$TMB, 0)` secs), 2. PCA-AGHQ (new, `r round(time_taken$aghq, 0)` hour), 3. NUTS (gold-standard, `r round(time_taken$tmbstan, 1)` days)
  * For PCA-AGHQ $k = 3$ and $s = 8$ chosen using Scree plot to explain ~90% of variance
  * For NUTS 4 chains of 100,000 thinned by 40 were required for good diagnostics e.g. all $\hat R < 1.025$

```{r fig4, echo=FALSE, fig.align='center', fig.cap='PCA-AGHQ moderately improves the posterior mean and substantially improves the posterior standard deviation'}
knitr::include_graphics("depends/mean-sd-alt-latent.png")
```

* PCA-AGHQ improves latent field point estimates (Figure 4) and distributional quantities like the Kolmogorov-Smirnov (KS, Figure 5) test statistic (`r ks_summary`% on average)

```{r fig5, echo=FALSE, fig.align='center', fig.cap='KS test demonstration.'}
knitr::include_graphics("fig5.png")
```

* Naomi can be used to assess probabilities targets have been met e.g. 90% of those who know their HIV status are on ART ("second 90")
  * Though PCA-AGHQ is better (`r round(second90_rmse, 0)`% RMSE), both TMB and PCA-AGHQ are biased (Figure 6)

```{r fig6, echo=FALSE, fig.align='center', fig.cap='PCA-AGHQ moderately improves the second 90 bias.'}
knitr::include_graphics("fig6.png")
```

@@

* Can we do better than moderate improvements?
  * Especially for the quantities of policy interest
* Laplace marginals with matrix algebra approximations [@wood2020simplified] to speed up calculations
* Further methods for allocation of effort to "important" dimensions of hyperparameter grid

**Contact**

* `r fontawesome::fa("code", fill = "black")` `github.com/athowes/naomi-aghq`
* `r fontawesome::fa("home", fill = "black")` `athowes.github.io`
* `r fontawesome::fa("envelope", fill = "black")` `ath19@ic.ac.uk`

@@
